---
title: "APS Coding Problem Set 1 ( Bats Problem )"
output:
  pdf_document: default
  html_document: default
date: "2023-08-22"
---


Load the CSV data file 'bats.csv' into the dataframe 'dfBat' and display the first few rows

```{r}
dfBat= read.csv('Data/bats.csv')
head(dfBat)
```



Display the structure of the dataframe 'dfBat' and its column names

```{r}
str(dfBat)
colnames(dfBat)
```



Create a vector 'gene_cols' containing the names of selected gene columns

```{r}
gene_cols = c("Gene.1", "Gene.2", "Gene.3", "Gene.4", "Gene.5", "Ebola")
```





**What is the chance of a random bat carrying the Ebola virus?**

```{r}
overall_ebola_chance = mean(dfBat$Ebola == TRUE)
ebola_chance = overall_ebola_chance*100
ebola_chance
```





**For each gene, calculate the likelihood that it is expressed in a random bat.**

```{r}
gene_likelihoods = sapply(gene_cols[1:5], function(col_name) {
  mean(dfBat[[col_name]]) * 100
})
gene_likelihoods
```





**Is the presence or absence of any of the genes indicative of a random bat potentially carrying the Ebola virus?**

```{r}
significant_genes = character(0) # Initialize an empty vector to store gene names

for (gene_col in gene_cols[1:5]) {
  gene_present_ebola_chance = mean(dfBat$Ebola[dfBat[[gene_col]]] == TRUE)
  gene_absent_ebola_chance = mean(dfBat$Ebola[!dfBat[[gene_col]]] == TRUE)

  if (abs(gene_present_ebola_chance - overall_ebola_chance) > 0.1 ||
      abs(gene_absent_ebola_chance - overall_ebola_chance) > 0.1) {
    significant_genes = c(significant_genes, gene_col)
  }
}

significant_genes
```